:::{.callout-note}
## Instruction

This is a quarto file. If you open its [source](report_template.qmd) in RStudio, you will get all the features of this interactive notebook-style document. If you want to know more about how to use this file to generate a .html document to hand in, look at [the getting started guide](https://quarto.org/docs/get-started/hello/rstudio.html).

(Delete this callout box before handing in your assignment.)
:::

```{r}
#| label: R packages
#| echo: false
#| warning: false
#| message: false

library(text2vec)
library(wordcloud)
library(magrittr)

# additional packages here
```

```{r}
#| label: data loading

data("movie_review")
```

# Data description

Describe the data and use a visualization to support your story. (approx. one or two paragraphs)

```{r}
#| label: eda visualization
#| warning: false

movie_review %$% wordcloud(review, 
                           min.freq = 10, 
                           max.words = 50, 
                           random.order = FALSE,
                           colors = brewer.pal(8, "Dark2"))


```
In the word cloud stop words like "the" and "also" are over represented, suggesting their removal would be beneficial as it would reduce dimentionality without impacting the clustering. That is because stop words are so frequent and carry relatively little meaning with them. Additionally, the words "movie" and "film" should also be removed for similar reasons. As all of the reviews concern this topic it is very likely they all mention either "movie" or "film" and thus, the words occurrence likely do not reflect any grouping within the reviews. Additionally, various forms of words "character" and "characters" 

```{r}
rows <- nrow(movie_review)
```

There is `r rows` reviews.

Histogram of the word count of reviews 

# Text pre-processing

Describe text pre-processing steps you have used (approx. one or two paragraphs)

# Text representaion

Briefly describe your text representation method. (approx. one or two paragraphs)

# Text clustering

Briefly describe which models you compare to perform clustering. (approx. two or three paragraphs)

# Evaluation & model comparison

Describe how you compare the methods and why. (approx. two or three paragraphs)


```{r}
#| label: table example
data.frame(
  model       = c("clustering model 1", "clustering model 2"),
  performance = c(1.2, 1.8),
  other       = c(0.5, 0.3),
  notes       = c("Some note", "another note")
)
```

Topic modeling may perform badly (ex overfitting) when used on short texts. Which is the case here. https://lazarinastoy.com/topic-modelling-limitations-short-text/ 

# Team member contributions

Write down what each team member contributed to the project.

- Author One: a, b, c
- Author Two: b, c, d
- Author Three: a, b, d
