
```{r}
#| label: R packages
#| echo: false
#| warning: false
#| message: false

library(text2vec)
library(wordcloud)
library(magrittr)
library(tidyverse)

# additional packages here
```

```{r}
#| label: data loading

data("movie_review")
```

# Data description

```{r}
#| label: eda histogram
#| warning: false

# Number of reviews
rows <- nrow(movie_review)

# TODO: Style histogram
movie_review %>% 
  mutate(length = str_count(review)) %>% 
  ggplot(aes(x = length)) +
  geom_histogram(bins = 50) +
  labs(x = "Number of words",
       y = "Frequency")
```

The data consists of `r rows` reviews of various movies. The histogram shows the distribution of their lengths. Most reviews seem to be on the shorter side, with less than 2500 words, but the distribution is majorly skewed. Upon further inspection, there seem to be a few data points that have such a big length, they don't show up on the histogram. One of these datapoints is the longest review, that has a length of 13708 words. Overall, reviews also don't appear to be shorter than roughly 1000 words. 

To start exploring the content of the reviews, a word cloud can be used to visualize the most common words in the corpus.

```{r}
#| label: eda wordcloud
#| warning: false

movie_review %$% wordcloud(review, 
                           min.freq = 10, 
                           max.words = 50, 
                           random.order = FALSE,
                           colors = brewer.pal(8, "Dark2"))


```

In the word cloud stop words like "the" and "also" are over represented, suggesting their removal would be beneficial as it would reduce dimensionality without impacting the clustering. That is because stop words are so frequent and carry relatively little meaning with them. Additionally, the words "movie" and "film" should also be removed for similar reasons. As all of the reviews concern this topic it is very likely they all mention either "movie" or "film" and thus, the words occurrence likely do not reflect any grouping within the reviews. Additionally, various forms of words "character" and "characters" 


# Text pre-processing

Describe text pre-processing steps you have used (approx. one or two paragraphs)

```{python}
#| label: pre-processing

from unidecode import unidecode
import simplemma
import string
import stopwords

""""

    Conversion to lower case;
    Removal of tokens with digits;
    Removal of tokens with length less than 3 characters;
    Removal of diacritical marks;
    Spelling normalization;
    Removal of stop words;
    Stemming (for English).

"""

# from: https://pythonspot.com/nltk-stop-words/
# import stopword from english
stops = stopwords.words('english')


def remove_interpunction_digits (text):
    # remove interpunction
    remove_digits = text.translate(str.maketrans('', '', string.punctuation))

    # remove digits
    result = remove_digits.translate(str.maketrans('', '', string.digits))
    return result


def lemmatisation (text_list):
    lemmatized_list = [simplemma.lemmatize(token, lang='en', greedy=True) for token in text_list]
    return lemmatized_list


def prepare_data(stopwords: list, sentence: str) -> str:
    # removal of punctuation and digits
    # todo: maybe not digit removal look at eta.
    sentence = remove_interpunction_digits(sentence).strip()

    # removal of diacritical marks (weird characteres like é)
    word_l_normalised = [unidecode(word) for word in sentence]

    # todo: >>insert spellchecking

    # lemmitise words
    word_lemmitised_l = lemmatisation(word_l_normalised)

    # stopword removal English
    word_l_tokenized = [token for token in word_lemmitised_l if token in stopwords]

    # join list into string with black spaces
    tokens_str = ' '.join(word_l_tokenized)

    return tokens_str
```


# Text representaion

Briefly describe your text representation method. (approx. one or two paragraphs)

# Text clustering

Briefly describe which models you compare to perform clustering. (approx. two or three paragraphs)

# Evaluation & model comparison

Describe how you compare the methods and why. (approx. two or three paragraphs)


```{r}
#| label: table example
data.frame(
  model       = c("clustering model 1", "clustering model 2"),
  performance = c(1.2, 1.8),
  other       = c(0.5, 0.3),
  notes       = c("Some note", "another note")
)
```

Topic modeling may perform badly (ex overfitting) when used on short texts. Which is the case here. https://lazarinastoy.com/topic-modelling-limitations-short-text/ 

# Team member contributions

Write down what each team member contributed to the project.

- Author One: a, b, c
- Author Two: b, c, d
- Author Three: a, b, d
