:::{.callout-note}
## Instruction

This is a quarto file. If you open its [source](report_template.qmd) in RStudio, you will get all the features of this interactive notebook-style document. If you want to know more about how to use this file to generate a .html document to hand in, look at [the getting started guide](https://quarto.org/docs/get-started/hello/rstudio.html).

(Delete this callout box before handing in your assignment.)
:::

```{r}
#| label: R packages
#| echo: false
#| warning: false
#| message: false

library(text2vec)
library(wordcloud)
library(magrittr)
library(tidyverse)

# additional packages here
```

```{r}
#| label: data loading

data("movie_review")

# your R code to load the data here
```

# Data description

```{r}
#| label: eda wordcloud
#| warning: false

# TODO: Style histogram
movie_review %>% 
  mutate(length = str_count(review)) %>% 
  ggplot(aes(x = length)) +
  geom_histogram(bins = 50) +
  labs(x = "Number of words",
       y = "Frequency")
```

The histogram shows the distribution of the review lengths. Most reviews seem to be on the shorter side, with less than 2500 words, but the distribution is majorly skewed. Upon further inspection, there seem to be a few data points that have such a big length, they don't show up on the histogram. One of these datapoints is the longest review, that has a length of 13708 words. Overall, reviews also don't appear to be shorter than roughly 1000 words. 

```{r}
#| label: eda wordcloud
#| warning: false

# your R code to generate the plot here

movie_review %$% wordcloud(review, 
                           min.freq = 10, 
                           max.words = 50, 
                           random.order = FALSE,
                           colors = brewer.pal(8, "Dark2"))

```

# Text pre-processing

Describe text pre-processing steps you have used (approx. one or two paragraphs)

# Text representaion

Briefly describe your text representation method. (approx. one or two paragraphs)

# Text clustering

Briefly describe which models you compare to perform clustering. (approx. two or three paragraphs)

# Evaluation & model comparison

Describe how you compare the methods and why. (approx. two or three paragraphs)


```{r}
#| label: table example
data.frame(
  model       = c("clustering model 1", "clustering model 2"),
  performance = c(1.2, 1.8),
  other       = c(0.5, 0.3),
  notes       = c("Some note", "another note")
)
```


# Team member contributions

Write down what each team member contributed to the project.

- Author One: a, b, c
- Author Two: b, c, d
- Author Three: a, b, d
